PANNEURO_analysis

### 2020-10-27 PANNEURO_CFS_merge_script_WHO-pres.Rmd
bmi added

```{r source file path}
#source raw data directory: data.raw_path
source("../PANCHANGE_raw_path.R")
```

@@

@@

```{r merge data frames}
dat <- left_join(
  coping_baseline,
  cfs_wave6, by = c("Sample", "ID")
)
```

>>>>

```{r source file path}
#source raw data directory: data.raw_path
source("/Users/alishpalmos/Documents/GitHub//PANCHANGE_raw_path.R")
```

@@

bmi <- read_rds(path = paste0(data_path, "/bmi.rds"))

@@

```{r merge data frames}
dat <- left_join(
  coping_baseline,
  cfs_wave6, by = c("Sample", "ID")
)

dat <- left_join(
  dat, bmi, by = c("Sample", "ID")
)
```

### 2020-10-27 PANNEURO_analysis_WHO_pres.Rmd
bmi added

```{r source file path}
#source data directory: data_path
source("../PANCHANGE_raw_path.R")
```

@@

## COVID probable diagnosis
```{r COVID probable diagnosis frequencies}
dat.raw %>%
  count(covid_probable_case_incl_screening, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Age
### Proportion tables
```{r age_category per Covid prop per total}
age_category_per_Covid_prop_total <- dat.raw %>%
  count(age_category, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2)) %>% 
  drop_na()

@@

```{r age_category per Covid prop per Covid}
age_category_per_Covid_prop_Covid <- dat.raw %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(age_category, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2)) %>% 
  drop_na()

kable(age_category_per_Covid_prop_Covid)
age_category_per_Covid_prop_Covid
```

### Cross tables
```{r age_category per Covid prop cross table}
age_category_per_Covid_prop_Covid_cross_table <- dat.raw %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(age_category, covid_probable_case_incl_screening) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = covid_probable_case_incl_screening, values_from = c(n, Prop))

kable(age_category_per_Covid_prop_Covid_cross_table)
age_category_per_Covid_prop_Covid_cross_table
```

### Collapse age categories
```{r Collapse age_category}
dat.raw <- dat.raw %>%
  mutate(age_category_collapsed_numeric =
           case_when(
             age_category_numeric == 1 ~ 0,
             age_category_numeric == 2 ~ 1,
             age_category_numeric == 3 ~ 2,
             age_category_numeric == 4 ~ 3,
             age_category_numeric == 5 ~ 4,
             age_category_numeric == 6 ~ 5,
             age_category_numeric == 7 ~ 6,
             age_category_numeric == 8 ~ 7,
             age_category_numeric == 9 ~ 8,
             age_category_numeric == 10 ~ 8,
             age_category_numeric == 11 ~ 8,
             age_category_numeric == 12 ~ 8,
             age_category_numeric == 13 ~ 8

           )
  )

dat.raw <- dat.raw %>%
  mutate(age_category_collapsed =
           recode_factor(
             age_category_collapsed_numeric,
             `0` = "16-18",
             `1` = "19-25",
             `2` = "26-35",
             `3` = "36-45",
             `4` = "46-55",
             `5` = "56-65",
             `6` = "66-70",
             `7` = "71-75",
             `8` = "76+"
           )
  )

dat.raw %>%
  freq(age_category_collapsed)
```

### Proportion tables
```{r age_category_collapsed per Sample prop per total}
age_category_collapsed_per_Covid_prop_total <- dat.raw %>%
  count(age_category_collapsed, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2)) %>% 
  drop_na()

kable(age_category_collapsed_per_Covid_prop_total)
age_category_collapsed_per_Covid_prop_total
```

```{r age_category_collapsed per Sample prop per Sample}
age_category_collapsed_per_Covid_prop_Covid <- dat.raw %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(age_category_collapsed, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2)) %>% 
  drop_na()

kable(age_category_collapsed_per_Covid_prop_Covid)
age_category_collapsed_per_Covid_prop_Covid
```

### Cross tables
```{r age_category_collapsed per Sample prop cross table}
age_category_collapsed_per_Covid_prop_Covid_cross_table <- dat.raw %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(age_category_collapsed, covid_probable_case_incl_screening) %>%
  drop_na() %>% 
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = covid_probable_case_incl_screening, values_from = c(n, Prop))

kable(age_category_collapsed_per_Covid_prop_Covid_cross_table)
age_category_collapsed_per_Covid_prop_Covid_cross_table
```

### Chi square test
```{r age_category per probable Covid chisquare test}
age_category_per_Covid_chissquare_test <- dat.raw %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  count(age_category_collapsed, covid_probable_case_incl_screening) %>%
  spread(key = covid_probable_case_incl_screening, value = n) %>%
  select(-age_category_collapsed) %>%
  chisq.test() %>%
  glance()

kable(age_category_per_Covid_chissquare_test)
age_category_per_Covid_chissquare_test
```

@@

## BMI not available

@@

###Chi-square

```{r ethnicity per probable Covid chisquare test}
ethnicity_per_Covid_chissquare_test <- dat.raw %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  count(Ethnicity_collapsed, covid_probable_case_incl_screening) %>%
  spread(key = covid_probable_case_incl_screening, value = n) %>%
  select(-Ethnicity_collapsed) %>%
  chisq.test() %>%
  glance()

kable(ethnicity_per_Covid_chissquare_test)
ethnicity_per_Covid_chissquare_test
```

# Neurological frequencies

```{r neurological frequencies alsfrs}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  select(
    starts_with("alsfrs"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

# Scoring for neuro measures

# ALSfrs
```{ALSfrs col names r}
als.col.names <- dat.raw %>%
  select(
    starts_with("alsfrs") & ends_with("numeric")
  ) %>%
  colnames()

>>>>

```{r source file path}
#source data directory: data_path
source("/Users/alishpalmos/Documents/GitHub//PANCHANGE_raw_path.R")
```

@@

## COVID probable diagnosis
```{r COVID probable diagnosis}
dat.raw %>%
  count(covid_probable_case_incl_screening, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

## Age
## Proportion tables
```{r age_category per Sample prop per total}
age_category_per_Covid_prop_total <- dat.raw %>%
  count(age_category, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r age_category per Sample prop per Sample}
age_category_per_Sample_prop_Sample <- dat.raw %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(age_category_per_Sample_prop_Sample)
age_category_per_Sample_prop_Sample
```

## Cross tables
```{r age_category per Sample prop cross table}
age_category_per_Sample_prop_Sample_cross_table <- dat.raw %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(age_category_per_Sample_prop_Sample_cross_table)
age_category_per_Sample_prop_Sample_cross_table
```

@@

```{r}

dat.raw$bmi[is.infinite(dat.raw$bmi)] = NA

dat.raw <- dat.raw %>%
  mutate(
  bmi.categories=cut(dat.raw$bmi, breaks=c(-Inf, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55,  Inf), labels=c("too low","10 and less","10 to 15", "15 to 20", "20 to 25", "25 to 30", "30 to 35", "35 to 40", "40 to 45", "45 to 50", "50 to 55")))

```{r bmi_category per Sample prop per total}

bmi_category_per_Covid_prop_total <- dat.raw %>%
  count(bmi.categories, covid_probable_case_incl_screening) %>%
  mutate(Prop = round(n/sum(n), 2))

options(knitr.kable.NA = '')
kable(bmi_category_per_Covid_prop_total)
bmi_category_per_Covid_prop_total
```
```{r bmi_category per Sample prop per Sample}
bmi_category_per_Sample_prop_Sample <- dat.raw %>%
  group_by(Sample) %>%
  count(bmi.categories, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(bmi_category_per_Sample_prop_Sample)
bmi_category_per_Sample_prop_Sample
```

@@

# Neurological frequencies
** NEED TO BE CHANGED TO covid_probable_case_incl_screening **
```{r neurological frequencies alsfrs}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  select(
    starts_with("alsfrs"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

**ADD SCORING FOR NEURO MEASURES**
# ALSfrs
```{ALSfrs col names r}
als.col.names <- dat.raw %>%
  select(
    starts_with("alsfrs") & ends_with("numeric")
  ) %>%
  colnames()
